---
title: "Economic Connections At Colleges"
author: Quinn Liu
output: 
  github_document:
    toc: true
    toc_depth: 5
---

```{r libraries, include = F}
library(knitr) 
library(downloader)
library(tidyverse)
```

## The Data

#### Description of the Data Sets

[Opportunity Insights](https://data.humdata.org/dataset/social-capital-atlas) has a data set on social capital, collecting data on friendships using privacy-protected data from Facebook. They explore the degree to which high-income and low-income people are friends with one another. They have used economic connectedness, cohesiveness, and civic engagement as measures of social capital and have collected data across counties, ZIP codes, high schools, and colleges. I have chosen to specifically explore colleges and economic connectedness. Economic connectedness (EC) is the share of high-income friends among low-income individuals.

I decided to combine this data set with another data set to combine social capital with other characteristics like private/public and admission rates to explore what characteristics are most correlated with higher economic connectedness. While the `rcis` library has the `scorecard` data set that provides the Department of Education’s College Scorecard data, I found and used the full spreadsheet of the DOE's [Scorecard data](https://collegescorecard.ed.gov/data/) to both obtain both more colleges and more variables that were not in the `rcis` data. This allowed me to retain most of the social capital data.

#### Retrieving and Cleaning the Data 

```{r import-data, include = F}
social <- read_csv(file = "data/social_capital_college.csv")

# download and unzip, file too big to push to git
download("https://ed-public-download.app.cloud.gov/downloads/Most-Recent-Cohorts-Institution_04192023.zip", 
         destfile = "data/scorecard.zip")
unzip("data/scorecard.zip", exdir = "data")
scorecard_doe <- read_csv(file = "data/Most-Recent-Cohorts-Institution.csv")

df <- inner_join(social, scorecard_doe, by = c("college_name" = "INSTNM"))
```

```{r clean-data, include = F}

types = factor("Public", "Private, non-profit", "Private, for-profit")

df <- df %>%                    # selecting variables to be used and of interest
  select(c(college_name, ec_own_ses_college, exposure_own_ses_college,  
           clustering_college, bias_own_ses_college, volunteering_rate_college, 
           CONTROL, HBCU, WOMENONLY, ADM_RATE, UGDS, COSTT4_A)) %>%
  rename(
    ec = ec_own_ses_college, 
    exposure = exposure_own_ses_college,
    type = CONTROL,
    adm_rate = ADM_RATE, 
    avg_cost = COSTT4_A, 
    is_hbcu = HBCU, 
    is_women_only = WOMENONLY
  ) %>%
  drop_na(ec) %>%
  mutate(                       
    ec = 0.5*ec,               # originally 2x share of high SES friends
    exposure = 0.5*exposure,   # originally 2x share of high SES individuals
    type = factor(type, labels = c("Public", "Private, non-profit", "Private, for-profit")), 
    adm_rate = as.double(adm_rate), 
    avg_cost = as.double(avg_cost), 
    is_hbcu = as.integer(is_hbcu), 
    is_women_only = as.integer(is_women_only)
  )
```
I retrieved the two data sets and joined them. The social capital data has 2586 observations, and the combined data set has 2571 observations, losing only 15 observations. And after cleaning out data for NA economic connectedness values, the data set used in general has 2351 observations.

```{r data, echo = F}
head(df)
```


#### Interest of the Data

From [Opportunity Insight's report and analysis by Chetty et al. (2022)](https://opportunityinsights.org/wp-content/uploads/2022/07/socialcapital_nontech.pdf), the data on social capital by county show that children from communities with higher economic connectedness are more likely to rise out of poverty. They obtained data on social capital by country and also followed up on data collection to obtain outcome measures to determine economic mobility. Outcome data was not obtained for colleges, but it can be weakly extrapolated that low-income students who have higher economic connectedness at college will have greater economic success. It is a common sentiment, particularly expressed by those involved in the college admission process, that one of the benefits of elite (often private) colleges is the networking and connections you are able to make. "It's not what you know, but who you know." And you get to know more rich and connected people at elite private colleges. The interest of this data set is to validate if indeed elite private colleges provide better economic connections and what are kind of characteristics high economic connectedness colleges have.
 
## Exploring the Data
 
**Main Question: What kind of college tends to have the highest economic connectedness? How do private and public colleges compare?**

#### Private vs Public Colleges
 
```{r ec-type-stat, include = F}
df %>% group_by(type) %>% count()

means = df %>%
  group_by(type) %>%
  summarize(ec = mean(ec))
```

```{r ec-type-histogram, echo = F}

df %>%
  ggplot(aes(x = ec, fill = type)) + 
  geom_histogram(bins = 40, aes(y = after_stat(density))) +
  facet_wrap(facets = vars(type)) + 
  geom_density(aes(x = ec), inherit.aes = F) +
  geom_vline(aes(xintercept = ec), means, linetype = "dashed") +
  annotate("text", label = "mean", x = 0.7, y = 6) +
  labs(
    title = "Distributions of Economic Connectedness Across Different Types of Colleges", 
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Economic Connectedness", 
    y = "Density"
  ) +
  scale_fill_manual(values = c("Public" = "skyblue3", 
                               "Private, non-profit" = "indianred2", 
                               "Private, for-profit" = "goldenrod")) +
  scale_x_continuous(labels = scales::percent) +  
  theme_gray() + 
  theme(legend.position = "none")
```

From the histograms and density curves, it can be seen that private, for-profit colleges have a greater density of colleges with higher economic connectedness. A greater percentage of private, non-profit colleges have high economic connectedness than the other two types. Private, non-profit colleges generally cluster around 80% with a heavy negative skew, while private, for-profit clusters around 65%. Public colleges appear to be bimodal, with its largest mode around 60% and a second mode around 75%. Private, non-profit colleges are the most likely type of college to have a high economic connectedness. 

The mean economic connectedness for private, non-profit is 80.3%, while it is 60.4% for public colleges and 63.4% for private, for-profit colleges. This tells us that on average, a low socioeconomic status student at a private, non-profit college has a greater percentage of high socioeconomic status friends than their other two counterparts. Indeed, it does seem that private colleges provide more economic connections,

Private, for-profit colleges' sparse distribution is due to a lack of observations. Private, for-profit colleges are less common than their counterparts, and in general, do not have the same regard in public sentiment. Of the 2351 observations, only 36 were private, for-profit. This limits the analysis possible for private, for-profit colleges, but we can still see that its distribution is closer to that of public colleges than private, non-profit colleges. 

##### Exposure

My conjecture for why private, non-profit colleges have the greatest proportion of high economic connectedness is that private colleges tend to cost more and thus have a greater proportion of high-income students. 

It can be seen that exposure is the best indicator of economic connectedness. Exposure is a variable from Opportunity Insight's Social Capital data set and is a measure of the share of high socioeconomic status individuals at a college.

```{r exposure-type, echo = F}

df %>%
  ggplot(aes(x = exposure, ec, color = type)) +
  geom_point(alpha = 0.25) + 
  facet_wrap(vars(type)) + 
  labs(
    title = "Exposure on Economic Connectedness Across Different Types of Colleges",
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Economic Connectedness",
    y = "Density"
  ) +
  scale_color_manual(values = c("Public" = "skyblue3",
                               "Private, non-profit" = "indianred2",
                               "Private, for-profit" = "goldenrod")) +
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(labels = scales::percent) +
  theme_gray() +
  theme(legend.position = "none")
```

There is a very strong linear relationship between exposure and economic connectedness. This occurs across public and private colleges. From these faceted scatterplots, it can also be seen that private, non-profit colleges have a greater cluster among high exposure and high economic connectedness, which is not seen in the other types. This aligns with the distribution that was seen in the economic connectedness histogram. This also supports the conjecture that private, non-profit colleges have higher economic connectedness due to a greater proportion of high-income students. Colleges with more high-income students allow the low-income students at these colleges to make more high-income friends due to the pure fact of who is around to befriend. 

Again, private, for-profit does not have a lot of data points. Since private, for-profit lacks sufficient data, here-on-forth, I will focus on public and private, non-profit colleges.

```{r type-clean, include = F}
# used in private vs public data here-on-forth
df_type <- df %>%
  filter(type != "Private, for-profit")
```

##### Friending Bias

Friending bias is a measure of how likely a low-income student is to be friends with the high-income students they are exposed to, calculated by dividing economic connected and exposure and then subtracting by 1.

```{r bias, echo = F}

df_type %>%
  ggplot(aes(x = bias_own_ses_college, fill = type)) +
  geom_histogram(bins = 40, aes(y = after_stat(density))) +
  facet_wrap(vars(type)) +
  geom_density(aes(x = bias_own_ses_college), inherit.aes = F) + 
    labs(
    title = "Distributions of Friending Bias",
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Friending Bias",
    y = "Density"
  ) +
  scale_color_manual(values = c("Public" = "royalblue",
                               "Private, non-profit" = "indianred2")) +
  scale_x_continuous(labels = scales::percent) + 
  theme_gray() +
  theme(legend.position = "none")
```

Across all types, most colleges have a friending bias around 0, meaning that students are not more likely to make friends with a high-income person than forming friendships with the general person at their college. And while both private and public colleges cluster around 0, private colleges have a greater density at 0, and public colleges have a slightly more spread out distribution. Public colleges seem to be slightly more likely to have a positive and a negative bias. General campus culture could be responsible for the slight differences in friending bias. 

##### Admission Rates

```{r adm-rate-type, echo = F, warning = F}

df %>%
  drop_na(adm_rate) %>%
  filter(type != "Private, for-profit") %>%
  ggplot(aes(x = adm_rate, y = ec, color = type)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(color = "honeydew4") + 
  facet_wrap(facets = vars(type)) + 
  labs(
    title = "College Admission Rate on Economic Connectedness", 
    subtitle = "Comparing Between Type of College", 
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Admission Rate", 
    y = "Economic Connectedness"
  ) +
  scale_color_manual(values =  c("Public" = "royalblue", 
                               "Private, non-profit" = "indianred2")) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  theme_gray() +
  theme(legend.position = "none")

```

We can see that across both public and private colleges, there is an overall negative correlation between the admission rate and economic connectedness of a college. The lower the admission rate, the higher the economic connectedness. We can see a small cluster of private colleges with very low admission rates. This cluster is not observable in public colleges, and public colleges do not reach the same minimum admission rate as private colleges. Comparing between private and public colleges, colleges of similar selectivity have a similar average economic connectedness. Instead, private colleges' association with higher economic connectedness seems to come from these very selective private colleges. I would conjecture that lower admission rates are correlated with economic connectivity because selective colleges often have a lot of resources and attract numerous applications, and richer individuals are more likely to be able to be admitted and attend. 

We can also explore the relationship between the average cost of attendance and EC. 

##### Average Cost of Attendance

```{r avg-cost-type, echo = F, warning = F}

df_type %>%
  drop_na(avg_cost) %>%
  ggplot(aes(x = avg_cost, y = ec, color = type)) +
  geom_point(alpha = 0.25) +
  geom_smooth(color = "azure4") +   
  facet_wrap(facets = vars(type)) + 
  labs(
    title = "College Average Cost of Attendance on Economic Connectedness", 
    subtitle = "Comparing Between Type of College", 
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Average Cost", 
    y = "Economic Connectedness"
  ) +
  scale_color_manual(values =  c("Public" = "royalblue", 
                               "Private, non-profit" = "indianred2")) +
  scale_y_continuous(labels = scales::percent) +
  theme_gray() +
  theme(legend.position = "none")

```

```{r avg-cost-mean, include = F}
df_type %>%
  group_by(type) %>%
  summarize(means = mean(avg_cost, na.rm = T))
```

There is an observable positive correlation between the average cost of attendance and EC in both private and public colleges. It can also be seen that public colleges do not exceed `$`40000, while the maximum average cost for private colleges is twice that. The mean of the average public college cost is 16898.78, while the mean for private college is around 49398.96. The average cost can be a measure of who can afford the college. While low-income students often get substantial financial aid, middle-class individuals often do not get enough financial aid to afford expensive colleges. Thus, colleges with high tuition can be a barrier to middle-class individuals and instead will admit more high-income students. The barrier for low-income students tends to be less cost and more a lack of resources to building an impressive college application. In general, average cost tells us the economic background of the average student, and the higher the cost, the more high-income individual there are, and the more economically connected a low-income student will be.

There is a correlation among all these factors: economic connectedness, exposure, admission rate, and average cost. It is not clear which of these, if any, causes the other factor, but a high economic connected college will likely be a selective, expensive, and private college. I would conjecture that exposure is likely the cause of economic connectedness of a college, as the relationship shown is very strong. Admission rate and average cost had a weaker correlation to EC, and their correlation likely occurs more through exposure.


#### Colleges with Focused Demographics

Some colleges are aimed at particular demographics, like a historically Black College and University (HBCU) or Women's Colleges. Since these aim to support a demographic that has been historically barred from education, it is interesting to see how economic connected these colleges are, along with other general statistics, in comparison to the general college. 

```{r college-means, include = F}

# function used for hbcu and women college to later give general statistics
college_means <- function(df, var){
  stats <- df %>%
    group_by({{var}}) %>%
    summarise(across(.col = where(is.double), .fns = mean, na.rm = T))
  stats <- t(stats[-1]) # transpose df w/o hbcu col
  rownames(stats) <- c("Economic Connectedness", "Exposure", "Cohesiveness",
                    "Friending Bias", "Volunteering Rate", "Admission Rate",
                    "Average Cost")
  return(stats)
}
```


##### Historically Black College and University

The DOE College Scorecard provides a variable for if a college is a historically Black College and University (HBCU). 

```{r hbcu-clean, include = F}
df_hbcu <- df %>%
  drop_na(is_hbcu)
```

```{r hbcu-count, include = F}
df_hbcu %>% group_by(is_hbcu) %>% count()
```

```{r ec-hbcu, echo = F}
df_hbcu %>%#9F97CC
  ggplot(aes(x = ec, fill = factor(is_hbcu))) + 
  geom_histogram(bins = 40, aes(y = after_stat(density))) +
  scale_fill_manual(values = c("0" = "#9F97CC", "1" = "#54428E")) +
  facet_wrap(facets = vars(is_hbcu), 
             labeller = labeller(is_hbcu = c("0" = "Not HBCU", "1" = "HBCU"))) + 
  geom_density(aes(x = ec), inherit.aes = F) + 
  labs(
    title = "Economic Connectedness of Historically Black Colleges and Universities", 
    subtitle = "Compared to the General College",
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Economic Connectedness", 
    y = "Density"
  ) +
  scale_x_continuous(labels = scales::percent) +
  theme_gray() +
  theme(legend.position = "none")
```

```{r hbcu-stats, echo = F}
college_means(df_hbcu, is_hbcu) %>%     
  kable(
    caption = "Mean of HBCU Colleges Characteristics", 
    col.names = c("Not HBCU", "HBCU"),
    digits = 5)
```

There are 87 HBCU colleges and 2256 non-HBCU colleges in the histogram. HBCU colleges are not as numerous, and thus, the distribution of HBCU colleges is more sparse than the distribution for non-HBCU colleges. However, the distribution do tell us that the EC of an average HBCU college is less than the EC of an average non-HBCU college. The minimum EC of any HBCU college is greater than that of any non-HBCU college, and the maximum EC of any HBCU is lower than that of any non-HBCU college. 

Since the target student population of HBCU colleges are black students, a larger proportion of high-incomes families are typically white, and thus, HBCU colleges have a lower proportion of high-income students. We can see from the table of means of college characteristics for HBCU colleges have on average lower exposure. However, the mean admission rate is slightly lower for HBCU colleges, with slightly lower average cost too. The comparison trends in private vs public colleges may not apply to HBCU colleges for admission rate and average cost, but it generally seems like HBCU colleges do not fit the typical high EC college.

##### Women's Colleges


```{r women-clean, include = F}
df_women <- df %>%
  drop_na(is_women_only)
```

```{r women-count, include = F}
df_women %>% group_by(is_women_only) %>% count()
```

```{r ec-women, echo = F}
df_women %>%
  ggplot(aes(x = ec, fill = factor(is_women_only))) + 
  geom_histogram(bins = 40, aes(y = after_stat(density))) +
  scale_fill_manual(values = c("0" = "lightsteelblue3", "1" = "steelblue4")) +
  facet_wrap(facets = vars(is_women_only), 
             labeller = labeller(is_hbcu = c("0" = "Not Women's College", "1" = "Women's College"))) + 
  geom_density(aes(x = ec), inherit.aes = F) + 
  labs(
    title = "Economic Connectedness of Women's Colleges", 
    subtitle = "Compared to the General College",
    caption = "Source: Opportunity Insight and DOE Scorecard",
    x = "Economic Connectedness", 
    y = "Density"
  ) +
  scale_x_continuous(labels = scales::percent) +
  theme_gray() +
  theme(legend.position = "none")

```

```{r women-stats, echo = F}
college_means(df_women, is_women_only) %>%     
  kable(
    caption = "Mean of Women's Colleges Characteristics", 
    col.names = c("Not Women's College", "Women's College"),
    digits = 5)
```

There are 27 Women's Colleges and 2316 non-Women's Colleges in the graphs, thus a sparse distribution for Women's Colleges, but from what is observable in the histograms, it can be seen that both Women's Colleges and non-Women's Colleges both have a negative skew and cluster more around a high EC. The EC of any Women's Colleges does not go below 50%. We can also see from the table of means of college characteristics that the average Women's College has more EC, more exposure, lower admission rates, and higher average costs. Women's Colleges fit the general trends for what a high EC college looks like. It seems that the general student population that Women's Colleges attract and admit is more likely to be high-income. 
 
It could be interesting for later further research to explore the subset of colleges that are both women-only and historically black, though the number of colleges that fit this category is likely to be very small.


## Summary

A high EC college is one with a large proportion of high-income students, and a college that admits a large proportion of high-income students is, on average, a selective, expensive, and private college. HBCU colleges are not typically a high EC college, while Women's Colleges are typically a selective, expensive, high EC college.

While Opportunity Insights found that high economic connectedness leads to greater economic mobility in counties, this does not mean that there is very strong reason to believe that high EC in college will lead to greater economic success. There is some general sense that being able to network with students that come from higher socioeconomic statuses, who themselves have better connections, will in some way lead to better success, the current data set has little way to provide evidence for such. I would be compelled to see further research and data collection into outcome results for college EC. 

